Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-06-19 edubots_surveys.Rnw
Advances the state of the art #6
Also adds a simple script for knitting

The rest of the paper is organized as follows. Next is an overview of what current research has found about the use of messaging applications, including chatbots, in the classroom. The methodology used in the survey is presented in Section \ref{sec:meth}, and the results of the survey are presented next in Section \ref{sec:res}. Finally, we discuss these results and conclude with a series of recommendations in the section that closes the paper.

@@

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant messaging (MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook Messenger stems from their simplicity, ease of use and multi-modality (i.e. video, audio, text) \cite{tang2017mobile}.  Using these tools in higher education can facilitate the delivery of personalised learning that occurs anytime anywhere, promote collaborative learning experiences and group discussions \cite{panah2020study}. WhatsApp is the most popular MIM platform used by educators to give assignments' feedback to students, support course discussions and provide learning resources in informal learning settings \cite{panah2020study}.  The use of WhatsApp in higher education could enhance social presence \cite{tang2017mobile} and foster trust relationships between educators and students embedded in the social learning process \cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due to the blurring of boundaries between academic and private life. This can lead to technostress \cite{gachago2015crossing}, difficulty in managing responsibilities especially among mature students and lack of privacy \cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave groups at any time can hinder their learning and undermine educators' efforts \cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and norms for these MIM groups in order to maintain the safety of these online communities for students \cite{abbas2021onlinechat} without affecting students' ownership and control that is vital to advance their learning \cite{gachago2015crossing}.

% The Use of Chatbots in Higher Education
The use of conversational agents or chatbots in higher education is still at its infancy \cite{yang2019opportunities}. Nevertheless, recent studies examining their positive impact on students' academic performance \cite{perez2020rediscovering} and engagement \cite{differ.chat}(Abbas et al., 2021b) have led to a growing interest in using this technology in classes. Indeed, the use of chatbots in collecting course feedback from students in higher education improved students' response quality and boosted their enjoyment levels \cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using either tools such as mobile devices or teaching strategies based on gamification \cite{Yildirim2017} can improve student motivation. In this sense, \cite{Pimmer2019} carried out a series of experiments with technological tools whose results showed positive perception and acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies were undertaken. For instance, an evaluation review study presented by \cite{Smutny2020} examined 47 educational chatbots implemented in Facebook Messenger with the focus to identify characteristics and quality metric such as language, subject matter and platform. Another study aimed to categorise educational chatbots according to their purpose into service-oriented and teaching-oriented \cite{perez2020rediscovering}. The first category includes those that provide service support such as the chatbot Ask Holly \cite{meetholly} and Dina \cite{santoso2018dinus}, both chatbots respond to students' questions about enrolment and registration. Ask L.U. \cite{LancasterUniversity_2019} answers students' frequently asked questions about timetables, grades, tutors and societies. LISA \cite{dibitonto2018chatbot} and Differ \cite{differchat} facilitate breaking the ice between new students by introducing them to each other.  \cite{ranoliya2017chatbot} proposed a generic chatbot for university students that is able to answer their frequently asked questions; for some time, the University of Granada had ELVIRA as a chatbot in their main page \cite{MOREO20129}. Besides being able to answer pre-established (frequent) questions, it did so from the main University of Granada using an inset character, voiced by a real person, who lip-synched the answers. The emphasis in this case, however, was on the authenticity of the synchronization more than in making the update of questions and answers simpler, or more interactive, using actual messaging platforms. In most cases, using the inset search engine was able to give you more up-to-date and accurate answers.

@@

Furthermore, \cite{sjostrom2018designing} proposed a conceptual architecture for teaching-oriented chatbots in higher education.  This conceptual architecture is based on a systematic literature review of previous studies examining the design of chatbots in higher education and a content analysis of student emails and discussion forum posts of four instances of a Java programming course.  The study outlined several design considerations, among them, the authors emphasised the importance of developing chatbots in platforms that students and educators are familiar with and can easily access (i.e. Facebook Messenger) which confirms with \cite{hobert2019you} and \cite{fernoagua2018intelligent} studies. In addition, \cite{sjostrom2018designing} argued that a conceptualision of learners' questions could aid designers in integrating the appropriate types of questions that the chatbots should support for different courses.  \cite{coronado2018cognitive} proposed agents that store learning materials to be provided on demand to students. \cite{crockett2017predicting} reported on tutoring systems, which can perform initial assessments of students' understanding and provide learning material that would advance their understanding to the next level.

>>>>

The rest of the paper is organized as follows: Next is an overview of what current research has found about the use of messaging applications, including chatbots, in the classroom. The methodology used in the survey is presented in Section \ref{sec:meth}, and the results of the survey are presented next in Section \ref{sec:res}. Finally, we discuss these results and conclude with a series of recommendations in the section that closes the paper.

@@

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education can facilitate the delivery of personalised learning
that occurs anytime anywhere, promote collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries,
the most popular MIM platform used by educators to give assignments'
feedback to students, support course discussions and provide learning
resources in informal learning settings \cite{panah2020study}.  The
use of WhatsApp in higher education could enhance social presence
\cite{tang2017mobile} and foster trust relationships between educators
and students embedded in the social learning process
\cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due
to the blurring of boundaries between academic and private life. This
can lead to technostress \cite{gachago2015crossing}, difficulty in
managing responsibilities, especially among mature students, and lack of
privacy \cite{tang2017mobile}.  Students' dropout of the MIM groups,
as they can leave groups at any time, can hinder their learning and
undermine educators' efforts \cite{mwakapina2016whatsapp}.  In
addition, there is a need to set rules and norms for these MIM groups
in order to maintain the safety of these online communities for
students \cite{abbas2021onlinechat} without affecting students'
ownership and control that is vital to advance their learning
\cite{gachago2015crossing}. This is why examining the role of MIM in higher education is still a challenge, and why the attitudes of the teaching community towards them has to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the use of chatbots, which, being conversational agents, usually well where synchronous conversations take place. 
The use of conversational agents or chatbots in higher education is
still at its infancy \cite{yang2019opportunities}. Nevertheless,
recent studies examining their positive impact on students' academic
performance \cite{perez2020rediscovering} and engagement
\cite{differ.chat,abbas2021onlinechat} have led to a growing interest
in using this technology in the (possibly virtual) classroom. Indeed, the use of chatbots in
collecting course feedback from students in higher education improved
students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using
either tools such as mobile devices or teaching strategies based on
gamification \cite{Yildirim2017} can improve student motivation. In
this sense, \cite{Pimmer2019} carried out a series of experiments with
technological tools whose results showed positive perception and
acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies were
undertaken. For instance, an evaluation review study presented by
\cite{Smutny2020} examined 47 educational chatbots implemented in
Facebook Messenger with the focus to identify characteristics and
quality metric such as language, subject matter and platform. Another
study aimed to categorise educational chatbots according to their
purpose into service-oriented and teaching-oriented
\cite{perez2020rediscovering}. The first category includes those that
provide service support such as the chatbot Ask Holly \cite{meetholly}
and Dina \cite{santoso2018dinus}, both chatbots respond to students'
questions about enrollment and registration. Ask
L.U. \cite{LancasterUniversity_2019} answers students' frequently
asked questions about timetables, grades, tutors and societies. LISA
\cite{dibitonto2018chatbot} and Differ \cite{differchat} facilitate
breaking the ice between new students by introducing them to each
other.  \cite{ranoliya2017chatbot} proposed a generic chatbot for
university students that is able to answer their frequently asked
questions; for some time, the University of Granada had ELVIRA as a
chatbot in their main page \cite{MOREO20129}. Besides being able to
answer pre-established (frequent) questions, it did so from the main
University of Granada using an inset character, voiced by a real
person, who lip-synched the answers. The emphasis in this case,
however, was on the authenticity of the synchronization more than in
making the update of questions and answers simpler, or more
interactive, using actual messaging platforms. In most cases, using
the inset search engine was able to give you more up-to-date and
accurate answers.

@@

Despite its (arguably) successful implementation in different higher education institutions, over all in the fringes of educational activity, its implementation or deployment is not trivial.
Some authors \cite{sjostrom2018designing} have proposed a conceptual
architecture for the adoption teaching-oriented chatbots in higher education.  This
conceptual architecture is based on a systematic literature review of
previous studies examining the design of chatbots in higher education
and a content analysis of student emails and discussion forum posts of
four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasised the
importance of developing chatbots in platforms that students and
educators are familiar with and can easily access (i.e. Facebook
Messenger) which confirms with \cite{hobert2019you} and
\cite{fernoagua2018intelligent} studies. In addition,
\cite{sjostrom2018designing} argued that a conceptualision of
learners' questions could aid designers in integrating the appropriate
types of questions that the chatbots should support for different
courses.  \cite{coronado2018cognitive} proposed agents that store
learning materials to be provided on demand to
students. \cite{crockett2017predicting} reported on tutoring systems,
which can perform initial assessments of students' understanding and
provide learning material that would advance their understanding to
the next level.
